---
title: "Practice 7.2"
output: html_notebook
---

# Visualization

## Load the needed libraries

```{r}
# Load the xUCINET library
library(xUCINET)

# Load statnet
library(statnet)
```

## Layout using ordination/scaling methods

Using the Wolfe_Primates network

For Multidimensional scaling you can either use the cmdscale function from statnet or the xMDS function from xUCINET

```{r}
# Using cmdscale function from stats library
WPJ_Coord <- cmdscale(max(Wolfe_Primates$JointPresence)-Wolfe_Primates$JointPresence+1, eig=TRUE, k = 2)
```

```{r}
# Using xMDS function from xUCINET library
WPJ_Coord <- xMDS(max(Wolfe_Primates$JointPresence)-Wolfe_Primates$JointPresence+1)
```

Plot

```{r}
par(mar = c(0, 0, 0, 0))
gplot(Wolfe_Primates$JointPresence>5,
      gmode="graph",
      coord=WPJ_Coord$points,
      vertex.cex=.8,
      vertex.col=(Wolfe_Primates$Attributes$Gender==1)*8,
      displaylabels=TRUE,
      label.pos=1,
      label.cex=.7,
      edge.col="grey70")
```

Using the Savage_TransactionFlows network

Recreate the Export Percentage symmetrized matrix from Practice 7.1

```{r}
Savage_TransactionFlows$Attributes[,4]<-Savage_TransactionFlows$Attributes$Longitude
Savage_TransactionFlows$Attributes[c(2,3),4]<-Savage_TransactionFlows$Attributes$Longitude[c(2,3)]+30
colnames(Savage_TransactionFlows$Attributes)[4]<-"Longitude_Adj"
Savage_TransactionFlows$Attributes

Savage_TransactionFlows$ExportPerc_SymSqrtProd<-(Savage_TransactionFlows$ExportPerc*t(Savage_TransactionFlows$ExportPerc))^.5
Savage_TransactionFlows$NetworkInfo<-rbind(Savage_TransactionFlows$NetworkInfo,c("ExportPerc_SymSqrtProd","A","FALSE","FALSE","Ratio","matrix"))
Savage_TransactionFlows$NetworkInfo
```

Multidimensional Scaling

```{r}
STE_Coord <- xMDS(max(Savage_TransactionFlows$ExportPerc_SymSqrtProd)-Savage_TransactionFlows$ExportPerc_SymSqrtProd+1)
```

Plot

```{r}
par(mar = c(0, 0, 0, 0))
gplot(Savage_TransactionFlows$ExportPerc_SymSqrtProd>5,
      gmode="graph",
      coord=STE_Coord$points,
      vertex.cex=.8,
      displaylabels=TRUE,
      label.pos=1,
      label.cex=.7,
      edge.col="grey70")
```

## Using MDS on geodesic distance matrix

Using the Advise network from practice 7.1

```{r}
KHA_Coord <- xMDS(geodist(xSymmetrize(Krackhardt_HighTech$Advice))$gdist)
```

plot

```{r}
par(mar = c(0, 0, 0, 0))
gplot(((Krackhardt_HighTech$Advice)*t(Krackhardt_HighTech$Advice)),
      gmode="graph",
      coord=KHA_Coord$points,
      vertex.cex=.8,
      displaylabels=TRUE, 
      label.pos=1,
      label.cex=.7,
      edge.col="grey70")
```
