---
title: "Practice 7.5"
output: html_notebook
---

# Visualization

## Load the required libraries

```{r}
library(xUCINET)
library(statnet)
```

## Representing directed networks

-   How to draw straight and curvy lines for directed networks using top three preferences among the members of Newcomb’s fraternity

Directed ties can be represented by straight lines in *sna*:

```{r}
par(mar=c(0,0,0,0))
gplot(Newcomb_Fraternity$PreferenceT00<4,
      gmode="digraph",
      #layout
      jitter=F,
      #ties
      edge.col="grey30",
      edge.lwd=1,
      arrowhead.cex=.6,
      #nodes
      vertex.col="gold", 
      vertex.cex=1.4,
      #labels
      displaylabels=TRUE, 
      label.pos=5,
      label.cex=.7)
```

Or curvy lines:

```{r}
par(mar=c(0,0,0,0))
gplot(Newcomb_Fraternity$PreferenceT00<4,
      gmode="digraph",
      #layout
      jitter = F,
      #ties
      edge.col="grey30",
      edge.lwd=1,
      usecurve=T,
      edge.curve=.09, 
      arrowhead.cex=.6,
      #nodes
      vertex.col="gold", 
      vertex.cex=1.4,
      #labels
      displaylabels=TRUE, 
      label.pos=5,
      label.cex=.7)
```

## Representing valued networks

-   How to include values to represent tie weights for edges as for the strength of trade among countries in 1928 depicted in Figure 7.7 and the interactions among a troop of monkeys in Figure 7.8.

We will not only make the tie width dependent on the values of the ties, but also the darkness.

```{r}
STE_1<-(Savage_TransactionFlows$ExportPerc*t(Savage_TransactionFlows$ExportPerc))^.5
maxV<-round(max(STE_1))+1
colfunc <- colorRampPalette(c("grey100", "grey5"))
COLWB<-colfunc(maxV)

STE_1Color<-STE_1
#create color for ties
for (k in (0:(maxV-1)))
{
  STE_1Color[STE_1>k]<-COLWB[k+1]
}

par(mar=c(0,0,0,0))
gplot((STE_1>4)*STE_1, 
      gmode="graph",
      coord=Savage_TransactionFlows$Attributes[,c(4,2)],
      edge.col=STE_1Color,
      edge.lwd=.3,
      vertex.col="orange",
      vertex.cex=.3,
      displaylabels=TRUE,
      label.pos=0,
      label.cex=.7)
```

Troops of Monkeys

```{r}
maxV<-max(Wolfe_Primates$JointPresence)+1
colfunc <- colorRampPalette(c("grey100", "grey5"))
COLWB<-colfunc(maxV)

WPJ_1Color<-Wolfe_Primates$JointPresence
#create color for ties
for (k in (0:(maxV-1)))
{
  WPJ_1Color[Wolfe_Primates$JointPresence>k]<-COLWB[k+1]
}

par(mar=c(0,0,0,0))
gplot((Wolfe_Primates$JointPresence-5)*(Wolfe_Primates$JointPresence>5), 
      gmode="graph",
      coord=Wolfe_Primates$Attributes[,2:3],
      edge.col=WPJ_1Color,
      edge.lwd=.3,
      vertex.col="orange",
      vertex.cex=.3,
      displaylabels=TRUE,
      label.pos=0,
      label.cex=.7)
```

For a directed network with values, it is generally better to use curved lines.

```{r}
par(mar=c(0,0,0,0))
gplot((Newcomb_Fraternity$PreferenceT00<4)*(4-Newcomb_Fraternity$PreferenceT00),
      gmode="digraph",
      #layout
      jitter=F,
      #ties
      edge.col="grey30",
      edge.lwd=1,
      usecurve=T,
      edge.curve=.09,
      #nodes
      vertex.col="gold", 
      vertex.cex=1.4,
      #labels
      displaylabels=TRUE, 
      label.pos=5,
      label.cex=.7)
```

## Using color and line style to represent multiple ties

-   How to add other characteristics of lines, such as color and line style (as for the games and antagonistic networks among the Hawthorne’s Bankwiring bank wiring employee in Figure 7.20)

To represent positive, negative and ambivalent ties, we use red, green and black lines,

```{r}
HBG<-Hawthorne_BankWiring$Game
HBA<-Hawthorne_BankWiring$Antagonistic

#Recode so that: only positive has color green (code3)
#only negative has color red (code 2 in R)
#both is black (code 1 in R)
HB<-(2*HBG+HBA)
HBc<-(HBA==1)*(HBG==1)*1+(HBA==0)*(HBG==1)*3+(HBA==1)*(HBG==0)*2
#Change tie width so it is also visible in black & white
HBl<-(HBA==1)*(HBG==1)*5+(HBA==0)*(HBG==1)*3+(HBA==1)*(HBG==0)*1


par(mar=c(0,0,0,0))
gplot(HB, jitter = F, gmode="graph", displaylabels=TRUE, label.pos=5, 
      vertex.col="grey90", edge.lwd=HBl,
      label.cex=.7, vertex.cex=1.4, edge.col=HBc)
```

```{r}
#alternative with dots
gplot(HB, jitter = F, gmode="graph", displaylabels=TRUE, label.pos=5, 
      vertex.col="grey90", edge.lwd=HBl, edge.lty=HBc,
      label.cex=.7, vertex.cex=1.4)
```
